### 2022-12-14 chapter1/chapter1.tex
Added missing_training figure to supplement

All HMM algorithms were implemented with custom code which is available as the package Homomorph on PyPI. The insertion phylo-HMM was trained on 39,298 manually labeled columns in 11 alignments, and the deletion phylo-HMM was trained on 58,655 manually labeled positions in 20 sequences in 9 unique alignments (Fig.~\ref{sfig:insertion_training}). Because maximum-likelihood estimation of the model parameters yielded posterior decoding curves which toggled between hidden states too rapidly, the models were instead trained discriminatively~\cite{Krogh1999}. The difference, briefly, is maximum-likelihood estimation finds the parameters that best reproduce the observed distributions whereas discriminative training finds the parameters that minimize prediction error. Discriminatively trained models typically perform better in practice since real-world data are rarely fully described by the distribution specified by the model.

>>>>

All HMM algorithms were implemented with custom code which is available as the package Homomorph on PyPI. The insertion phylo-HMM was trained on 39,298 manually labeled columns in 11 alignments, and the deletion phylo-HMM was trained on 58,655 manually labeled positions in 20 sequences in 9 unique alignments (Fig.~\ref{sfig:insertion_training}, Fig.~\ref{sfig:missing_training}). Because maximum-likelihood estimation of the model parameters yielded posterior decoding curves which toggled between hidden states too rapidly, the models were instead trained discriminatively~\cite{Krogh1999}. The difference, briefly, is maximum-likelihood estimation finds the parameters that best reproduce the observed distributions whereas discriminative training finds the parameters that minimize prediction error. Discriminatively trained models typically perform better in practice since real-world data are rarely fully described by the distribution specified by the model.

### 2022-12-14 chapter1/chapter1_supporting.tex
Updated figure labels to match file names
Added missing_training figure to supplement

\begin{figure}[h!]
\includegraphics[width=\textwidth]{insertion_training/out/insertion_training.png}
\centering
\caption{\textbf{Insertion phylo-HMM data and training details.}
\textbf{(A)} Most columns in the training data were labeled as state 1A or 1B. \textbf{(B)} The model loss stabilized by the final training iteration. \textbf{(C-F)} The values of parameters in the phylogenetic process, the jump process, the pattern stickiness model, and the transition matrix, respectively, at each training iteration. The transition matrix plots are the transition rates to the state indicated on the vertical axis and given in log scale. Self transitions are excluded.}
\label{sfig:hmm_training}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{insertion_trim/out/insertion_trim.png}
\centering
\caption{\textbf{Insertion phylo-HMM trimming details.}
\textbf{(A)} Most alignments were not trimmed. Of the alignments with trims, most were trimmed only at the level of sequences. \textbf{(B)} Most trimmed regions were inferred primarily as state 2. \textbf{(C)} Most alignments with sequence trims have fewer than 10 segments removed. \textbf{(D)} Most alignments with region trims have fewer than five regions removed. \textbf{(E, G)} The number of non-gap symbols in sequence trims can vary considerably, but for nearly all sequence trims each non-gap symbol in the removed segment is aligned to fewer than five non-gap symbols on average. Only the lower 95\% of the distribution of the number of non-gap symbols in the sequence trims is shown. \textbf{(F, H)} The length of region trims can also vary considerably, but generally each region trims accounts for fewer than 10\% of the columns in the original alignment.}
\label{sfig:hmm_trim}
\end{figure}

>>>>

\begin{figure}[h!]
\includegraphics[width=\textwidth]{insertion_training/out/insertion_training.png}
\centering
\caption{\textbf{Insertion phylo-HMM data and training details.}
\textbf{(A)} Most columns in the training data were labeled as state 1A or 1B. \textbf{(B)} The model loss stabilized by the final training iteration. \textbf{(C-F)} The values of parameters in the phylogenetic process, the jump process, the pattern stickiness model, and the transition matrix, respectively, at each training iteration. The transition matrix plots are the transition rates to the state indicated on the vertical axis and given in log scale. Self transitions are excluded.}
\label{sfig:insertion_training}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{insertion_trim/out/insertion_trim.png}
\centering
\caption{\textbf{Insertion phylo-HMM trimming details.}
\textbf{(A)} Most alignments were not trimmed. Of the alignments with trims, most were trimmed only at the level of sequences. \textbf{(B)} Most trimmed regions were inferred primarily as state 2. \textbf{(C)} Most alignments with sequence trims have fewer than 10 segments removed. \textbf{(D)} Most alignments with region trims have fewer than five regions removed. \textbf{(E, G)} The number of non-gap symbols in sequence trims can vary considerably, but for nearly all sequence trims each non-gap symbol in the removed segment is aligned to fewer than five non-gap symbols on average. Only the lower 95\% of the distribution of the number of non-gap symbols in the sequence trims is shown. \textbf{(F, H)} The length of region trims can also vary considerably, but generally each region trims accounts for fewer than 10\% of the columns in the original alignment.}
\label{sfig:insertion_trim}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=\textwidth]{missing_training/out/missing_training.png}
\centering
\caption{\textbf{Missing phylo-HMM data and training details.}
\textbf{(A)} Most columns in the training data were labeled as state 1, which is referred to as the ``not missing'' state in the main text. \textbf{(B)} The model loss stabilized by the final training iteration. \textbf{(C-F)} The values of parameters in the phylogenetic process, the jump process, and the transition matrix, respectively, at each training iteration. The transition matrix plots are the transition rates to the state indicated on the vertical axis and given in log scale. Self transitions are excluded.}
\label{sfig:missing_training}
\end{figure}
